---
title: "CMR Results"
author: "Limoilou"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  bookdown::word_document2:
  citation_package: natbib
  keep_tex: yes
  fig_caption: yes
  latex_engine: pdflatex
bibliography: cmr.bib
csl: canadian-journal-of-zoology.csl
fontsize: 12pt
fontfamily: times
---

```{r setup, include=FALSE, cache = TRUE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, verbose=FALSE, cache=TRUE)
# knitr::opts_knit$set(root.dir = '~/Users/LimoilouARenaud/Documents/PostDocI/Projects/cmrSeal')
```

```{r include=FALSE}
library(knitr)
library(magrittr) # problems with tidyverse. Installed this instead
library(dplyr)
library(coda)
library(bookdown)
library(ggplot2)
library(ggthemes)
library(pander)
library(cowplot)
library(lubridate)
library(coda)
library(boot)
library(purrr)
```

# Results

```{r sample size}
library(here)
load(here("output.nosync/data/2022-05-13_resultsCMR.RData"))
data <- pup.data35
nlevels(droplevels(unique(data$myID)))
length(unique(data$myID)) #1371
length(unique(data$date)) #  308
table(data$sex)
# female   male     NA
#   1120   1253      8 #
t = table(data$myID)
#View(t)
mean(t)
table(t)
# 0   1   2   3   4   5   6   7   8
# 13 757 376 140  63  17  12   5   1
sum(t)
757/1371# 0.5521517
(376 + 140 +  63  + 17  +12  +  5  +  1)/1371 # 0.4478483

# make table 1
data$myID <- droplevels(data$myID)
t1 <- table(data$myID)
# View(t1)
tmp=data[data$myID %in% names(t1)[t1>1],]
#View(tmp)
nlevels(droplevels(unique(tmp$myID))) # 615 individus 615/1371
t2=table(tmp$myID)
#View(t2)

s.sizes.b = data %>%
  filter(mySite=="bic") %>%
  group_by(year) %>%
  mutate(date =ymd(date)) %>%
           summarise(min = min(date),
                     max= max(date),
                     unique.ID = length(unique(myID)),
                     nb.male=length(na.omit(unique(ifelse(sex=='male',myID,NA)))),
            nb.female=length(na.omit(unique(ifelse(sex=='female',myID,NA))))
)

recap.b=pup.data35 %>%
  filter(mySite=="bic") %>%
  group_by(year) %>%
  droplevels() %>%
  summarise(unique.ID = length(unique(myID)))
#write.csv2(recap.b, "output/data/recapture_b.csv")

# add prop of recapt for table 1
t1 <- table(pup.data35$myID)
tmp=pup.data35[pup.data35$myID %in% names(t1)[t1>1],]

t2=tmp %>%
  filter(mySite=="bic") %>% group_by(year) %>%
  droplevels %>%
  summarise(recapt = length(unique(myID)))
recap.b=cbind(recap.b, t2[,2])
recap.b=recap.b%>% mutate(prop=recapt/unique.ID)

s.sizes.b <- cbind(s.sizes.b, recap.b[,4])


s.sizes.m = data %>%
  filter(mySite=="metis") %>%
  group_by(year) %>%
  mutate(date =ymd(date)) %>%
           summarise(min = min(date),
                     max= max(date),
                     unique.ID = length(unique(myID)),
                     nb.male=length(na.omit(unique(ifelse(sex=='male',myID,NA)))),
            nb.female=length(na.omit(unique(ifelse(sex=='female',myID,NA))))
)

recap.m=pup.data35 %>%
  filter(mySite=="metis") %>%
  group_by(year,.drop = FALSE) %>%
  #droplevels() %>%
  summarise(unique.ID = length(unique(myID)))
#write.csv2(recap.m, "output/data/recapture_m.csv")

t1 <- table(pup.data35$myID)
tmp=pup.data35[pup.data35$myID %in% names(t1)[t1>1],]

t2=tmp %>%
  filter(mySite=="metis") %>% group_by(year,.drop = FALSE) %>%
  #droplevels() %>%
  summarise(recapt = length(unique(myID)))
recap.m=cbind(recap.m, t2[,2])
recap.m=recap.m%>% mutate(prop=recapt/unique.ID)

s.sizes.m <- cbind(s.sizes.m, recap.m[,4])


# here the birth dates as estimated from the models
meanBd_bic <-  meanBd_bic %>% mutate(across(where(is.numeric), round, 2))
meanBd_metis <-  meanBd_metis %>% mutate(across(where(is.numeric), round, 2))
```

From 1998 to 2019, `r nlevels(droplevels(unique(data$myID)))` pups were captured in the study area and included in analyses (`r sum(s.sizes.m$nb.female) + sum(s.sizes.b$nb.female)` females,  `r sum(s.sizes.m$nb.male) + sum(s.sizes.b$nb.male)` males, `r sum(data$sex=='NA', na.rm=T)` pups of unknown sex), for a total of `r nrow(data)` observations over `r length(unique(data$date))` sampling occasions. Table \@ref(tab:t1) describes the sample size per site and year. Individuals were captured on average 1.72 times, ranging from 1 (N = 757) to 8 (N = 1) captures. Captures occurred between May 20 to July 10. 

```{r import RData from R}
load(here::here("output.nosync/data/2022-05-11_resultsCMR.RData"))

aicT[1,2]<-"$p(.)\\phi(.)b(t)$"# could put math in markdown - tex though
aicT[2,2]<-"$p(t)\\phi(.)b(t)$"
aicT[3,2]<-"$p(t)\\phi(t)b(t)$"
aicT[4,2]<-"$p(.)\\phi(t)b(t)$"
aicT[5,2]<-"$p(1|t)\\phi(1|t)b(t)$"
#aicT[6,2]<-"$p()\\phi()\\gamma(t)$" # will have to ad 6 and 7

aicT=aicT %>% mutate(across(where(is.numeric), round, 2))

long_N_dt_b <- cbind(aicb1[,c(1,3)], aicb2[3], aicb3[3],aicb4[3], aicb5[3])
names(long_N_dt_b)<-c("year", "model1", "model2", "model3", "model4", "model5")
long_N_dt_b <- long_N_dt_b %>% mutate(across(where(is.numeric), round, 2))

bestmod <- apply(long_N_dt_b[,-1],1,which.min)
tmp <-cbind(long_N_dt_b$year,colnames(long_N_dt_b)[bestmod+1])
knitr::kable(tmp,caption = "Selected model per year, Bic, Qc, Canada.", "simple")

# Métis
aicTM[1,2]<-"$p(.)\\phi(.)b(t)$"# could put math in markdown - tex though
aicTM[2,2]<-"$p(t)\\phi(.)b(t)$"
aicTM[3,2]<-"$p(t)\\phi(t)b(t)$"
aicTM[4,2]<-"$p(.)\\phi(t)b(t)$"
aicTM[5,2]<-"$p(1|t)\\phi(1|t)b(t)$"
#aicTM[6,2]<-"$p()\\phi()\\gamma(t)$"

aicTM=aicTM %>% mutate(across(where(is.numeric), round, 2))

#knitr::kable(aicTM, caption = "Average WAIC per model over years, Métis, Qc, Canada.", align = "llr")

long_N_dt_m <- cbind(aicm1[,c(1,3)], aicm2[3], aicm3[3],aicm4[3], aicm5[3])
names(long_N_dt_m)<-c("year", "model1", "model2", "model3", "model4", "model5")
long_N_dt_m <- long_N_dt_m %>% mutate(across(where(is.numeric), round, 2))

bestmod <- apply(long_N_dt_m[,-1],1,which.min)
tmp <-cbind(long_N_dt_m$year,colnames(long_N_dt_m)[bestmod+1])
knitr::kable(tmp,caption = "Selected model per year, Métis, Qc, Canada.", "simple")


```
```{r best CJS model}
# give proper names to the models
waic.table[1,2]<-"$p(.)\\phi(.)$"# could put math in markdown - tex though
waic.table[2,2]<-"$p(t)\\phi(.)$"
waic.table[3,2]<-"$p(t)\\phi(1|t)$"
waic.table[4,2]<-"$p(.)\\phi(1|t)$"
waic.table[5,2]<-"$p(1|i)\\phi(.)$"
waic.table[6,2]<-"$p(1|i)\\phi(1|t)$" # will have to ad 6 and 7
waic.table[7,2]<-"$p(.)\\phi(.)$"
waic.table[8,2]<-"$p(t)\\phi(.)$"
waic.table[9,2]<-"$p(t)\\phi(1|t)$"
waic.table[10,2]<-"$p(.)\\phi(1|t)$"
waic.table[11,2]<-"$p(1|i)\\phi(.)$"
waic.table[12,2]<-"$p(1|i)\\phi(1|t)$"
waic.table <- waic.table %>% mutate(across(where(is.numeric), round, 2))

long_Surv_dt_b <- cbind(WAIC.b.1[,c(2,1)], WAIC.b.2[1], WAIC.b.3[1],WAIC.b.4[1], WAIC.b.5[1], WAIC.b.6[1])
names(long_Surv_dt_b)<-c("year", "model1", "model2", "model3", "model4", "model5", 'model6')
long_Surv_dt_b <- long_Surv_dt_b %>% mutate(across(where(is.numeric), round, 2))

bestmod <- apply(long_Surv_dt_b[,-1],1,which.min)
tmp <-cbind(long_Surv_dt_b$year,colnames(long_Surv_dt_b)[bestmod+1])
knitr::kable(tmp,caption = "Selected model per year, Bic, Qc, Canada.", "simple")


long_Surv_dt_m <- cbind(WAIC.m.1[,c(2,1)], WAIC.m.2[1], WAIC.m.3[1],WAIC.m.4[1], WAIC.m.5[1], WAIC.m.6[1])
names(long_Surv_dt_m)<-c("year", "model1", "model2", "model3", "model4", "model5", "model6")
long_Surv_dt_m <- long_Surv_dt_m %>% mutate(across(where(is.numeric), round, 2))

bestmod <- apply(long_Surv_dt_m[,-1],1,which.min)
tmp <-cbind(long_Surv_dt_m$year,colnames(long_Surv_dt_m)[bestmod+1])
knitr::kable(tmp,caption = "Selected model per year, Métis, Qc, Canada.", "simple")
```



We found convergence issues in years with low capture rates or sample sizes: 2013 (N=13 individuals, no recapture, Métis) and 2003 (N=29, recapture rate of 0.14, Bic Island). Otherwise, convergence was reached for all models and all years based on the Gelman-Rubin statistics (Gelman and Rubin 1992), with maximum values for superpopulation size of 1.2 for abundance and survival probability. In Bic, a Jolly-Seber model for abundance including a time effect on both *φ* and *p*, was selected based on the averaged WAIC over years (Table S\@ref(tab:ts1), with model 3 selected in 14 of the 16 years (`r 14/16*100` % of time, Table S\@ref(tab:ts2)). In Métis, model 3 was selected in 13 of the 16 years of study and was thus chosen as the most realistic model `r 13/16*100` % of time. In Bic, a Cormack-Jolly-Seber including a constant survival and a fixed time effect on the capture probability was selected in 8 of the 16 years; the WAIC was however lower than any other model (`r waic.table %>% filter(site=='bic') %>% slice(which.min(WAIC)) %>% pull(WAIC)`)(Table S\@ref(tab:ts3). In Metis, WAIC was similar across CJS models  but slightly lower (`r waic.table %>% filter(site=='metis') %>% slice(which.min(WAIC)) %>% pull(WAIC)`) for a model including a random time effect on survival and a fixed time effect on capture probability (Table S\@ref(tab:ts4). **Based on these 'selected' models, capture probabilities in the JS (Figure **SX**) and CJS (Figure XX) models are illustrated for a year with a low (panels a) and high (panels b) sample size for both model types.**
Mean birth dates, as estimated from the Cormack-Jolly-Seber model for pre-weaning survival, ranged from `r  meanBd_bic %>% slice(which.min(Mean)) %>% pull(Mean)` [`r  meanBd_bic %>% slice(which.min(Mean)) %>% pull(CIL)`,`r  meanBd_bic %>% slice(which.min(Mean)) %>% pull(CIH)`] to `r meanBd_bic %>% slice(which.max(Mean)) %>% pull(Mean)` [ `r meanBd_bic %>% slice(which.max(Mean)) %>% pull(CIL)`, `r meanBd_bic %>% slice(which.max(Mean)) %>% pull(CIH)`] at Bic Island and from `r  meanBd_metis%>% slice(which.min(Mean)) %>% pull(Mean)` [ `r  meanBd_metis %>% slice(which.min(Mean)) %>% pull(CIL)`, `r  meanBd_metis %>% slice(which.min(Mean)) %>% pull(CIH)`] to `r  meanBd_metis %>% slice(which.max(Mean)) %>% pull(Mean)` [ `r  meanBd_metis %>% slice(which.max(Mean)) %>% pull(CIL)`, `r  meanBd_metis %>% slice(which.max(Mean)) %>% pull(CIH)`] at Metis (Table S\@ref(tab:ts5)).

```{r abundance and survival}
results_all <- results_all %>% mutate(across(where(is.numeric), round, 2))
results_weanSurv <- results_weanSurv %>% mutate(across(where(is.numeric), round, 2))

# compare to gab's extract
weanSurvOut_m <-  weanSurvOut_m%>% mutate(across(where(is.numeric), round, 2))
weanSurvOut_m_summa <-  weanSurvOut_m_summa%>% mutate(across(where(is.numeric), round, 2))

summary(lm(weanSurv_mean~yr, data=weanSurvOut_m_summa))

weanSurvOut_b <- weanSurvOut_b%>% mutate(across(where(is.numeric), round, 2))
weanSurvOut_b_summa <-  weanSurvOut_b_summa%>% mutate(across(where(is.numeric), round, 2))

summary(lm(weanSurv_mean~yr, data=weanSurvOut_b_summa))
```

Overall, population size was higher at Bic than at Métis (Figure \@ref(fig:f1)). Mean pup abundance at Bic Island was `r results_all%>% filter(site=='bic') %>% summarise(mean = mean(N, na.rm=T)) %>% pull(mean) %>% round(2)` and `r results_all%>% filter(site=='metis') %>% summarise(mean = mean(N, na.rm=T)) %>% pull(mean) %>% round(2)` at Métis. Pup abundance varied from a minimum value of `r results_all %>% filter(site =='bic') %>% slice_min(order_by = N) %>% pull(N)` [`r results_all %>% filter(site =='bic') %>% slice_min(order_by = N) %>% pull(ymin)`, `r results_all %>% filter(site =='bic') %>% slice_min(order_by = N) %>% pull(ymax)`] pups in `r results_all %>% filter(site =='bic') %>% slice_min(order_by = N) %>% pull(yr)` to a maximum of `r results_all %>% filter(site =='bic') %>% slice_max(order_by = N) %>% pull(N)` [`r results_all %>% filter(yr == '2019' & site =='bic') %>% pull(ymin)`, `r results_all %>% filter(yr == '2019' & site =='bic') %>% pull(ymax)`] in `r results_all %>% filter(site =='bic') %>% slice_max(order_by = N) %>% pull(yr)` at Bic Island. At Métis, pup abundance varied from a minimum of `r results_all %>% filter(site =='metis') %>% slice_min(order_by = N) %>% pull(N)` [`r results_all %>% filter(site =='metis') %>% slice_min(order_by = N) %>% pull(ymin)`, `r results_all %>% filter(site =='metis') %>% slice_min(order_by = N) %>% pull(ymax)`] in `r results_all %>% filter(site =='metis') %>% slice_min(order_by = N) %>% pull(yr)` to a maximum of `r results_all %>% filter(site =='metis') %>% slice_max(order_by = N) %>% pull(N)` [`r results_all %>% filter(site =='metis') %>% slice_max(order_by = N) %>% pull(ymin)`, `r results_all %>% filter(site =='metis') %>% slice_max(order_by = N) %>% pull(ymax)`] pups in `r results_all %>% filter(site =='metis') %>% slice_max(order_by = N) %>% pull(yr)`. 

Pre-weaning survival was also higher at Bic Island and less variable than at Métis, but variability around mean estimates was generally high and highly dependant on capture effort (Figure \@ref(fig:f2)). Mean pre-weaning survival at Bic Island was `r weanSurvOut_b_summa %>% summarise(mean = mean(weanSurv_mean, na.rm=T)) %>% pull(mean) %>% round(2)` and `r weanSurvOut_m_summa %>% summarise(mean = mean(weanSurv_mean, na.rm=T)) %>% pull(mean) %>% round(2)` at Métis. The survival probability ranged from a low value of `r weanSurvOut_b_summa %>%slice_min(order_by = weanSurv_mean) %>% pull(weanSurv_mean) %>% nth(1)` [`r weanSurvOut_b_summa %>%slice_min(order_by = weanSurv_mean) %>% pull(weanSurv_cil) %>% nth(1)`, `r weanSurvOut_b_summa %>%slice_min(order_by = weanSurv_mean) %>% pull(weanSurv_cih)%>% nth(1)`] in `r weanSurvOut_b_summa%>%slice_min(order_by = weanSurv_mean) %>% pull(yr)%>% nth(1)` to a maximum value of `r weanSurvOut_b_summa %>%slice_max(order_by = weanSurv_mean) %>% pull(weanSurv_mean)` [`r weanSurvOut_b_summa %>%slice_max(order_by = weanSurv_mean) %>% pull(weanSurv_cil)`, `r weanSurvOut_b_summa %>%slice_max(order_by = weanSurv_mean) %>% pull(weanSurv_cih)`] in `r weanSurvOut_b_summa %>% slice_max(order_by = weanSurv_mean) %>% pull(yr)` at Bic. At Métis, pre-weaning survival varied from `r weanSurvOut_m_summa %>%slice_min(order_by = weanSurv_mean) %>% pull(weanSurv_mean) %>% nth(1)` [`r weanSurvOut_m_summa%>%slice_min(order_by = weanSurv_mean) %>% pull(weanSurv_cil)%>% nth(1)`, `r weanSurvOut_m_summa %>%slice_min(order_by = weanSurv_mean) %>% pull(weanSurv_cih)%>% nth(1)`] in `r weanSurvOut_m_summa %>% slice_min(order_by = weanSurv_mean) %>% pull(yr)%>% nth(1)` to a maximum value of `r weanSurvOut_m_summa%>%slice_max(order_by = weanSurv_mean) %>% pull(weanSurv_mean)` [`r weanSurvOut_m_summa %>%slice_max(order_by = weanSurv_mean) %>% pull(weanSurv_cil)`, `r weanSurvOut_m_summa %>% slice_max(order_by = weanSurv_mean) %>% pull(weanSurv_cih)`] in `r weanSurvOut_m_summa %>%slice_max(order_by = weanSurv_mean) %>% pull(yr)`. 


```{r}

  correlation_bic<- lbda_bic.3 %>% group_by(it) %>% 
    summarise(cor_sN=cor(weanSurv,Nsuper),
              cor_sgrowth=cor(weanSurv,growth))
  
quantile(correlation_bic$cor_sN,probs = c(0.025,.5,0.975))
  # 2.5%          50%        97.5% 
  # -0.372452859 -0.003792789  0.426867806 
hist(correlation_bic$cor_sN);abline(v=0,col='red')
mean(correlation_bic$cor_sN>0)  # pseudo p-value 0.4916667
  
quantile(correlation_bic$cor_sgrowth,probs = c(0.025,.5,0.975))
hist(correlation_bic$cor_sgrowth);abline(v=0,col='red')
mean(correlation_bic$cor_sgrowth>0)  # pseudo p-value
  
  
  
correlation_Metis<- lbda_metis.3 %>% group_by(it) %>% 
  summarise(cor_sN=cor(weanSurv,Nsuper),
            cor_sgrowth=cor(weanSurv,growth))
quantile(correlation_Metis$cor_sN,probs = c(0.025,.5,0.975))
# 2.5%        50%      97.5% 
# -0.5194995 -0.1353982  0.3595358 
hist(correlation_Metis$cor_sN);abline(v=0,col='red')
mean(correlation_Metis$cor_sN>0)  # pseudo p-value 0.2903333

round(quantile(correlation_Metis$cor_sgrowth,probs = c(0.025,.5,0.975)),2)
# 2.5%        50%      97.5% 
# -0.4441526 -0.1190146  0.4316046 
hist(correlation_Metis$cor_sgrowth);abline(v=0,col='red')
mean(correlation_Metis$cor_sgrowth>0)  # pseudo p-value 0.3186667

```

Finally, we investigated if pre-weaning survival was correlated either with pup abundance or the 'observed population growth rate', calculated as Nt+1/Nt, where N was pup abundance. Pre-weaning survival was not correlated with neither pup abundance at time t nor pup growth rate at Bic (median correlation with N [95% CRI]: `r  round(quantile(correlation_bic$cor_sN,probs = c(.5)), 2) ` [`r  round(quantile(correlation_bic$cor_sN,probs = c(.025)), 2) `, `r  round(quantile(correlation_bic$cor_sN,probs = c(.975)), 2) `], median correlation with growth rate [95% CRI]: `r  round(quantile(correlation_bic$cor_sgrowth,probs = c(.5)), 2) ` [`r  round(quantile(correlation_bic$cor_sgrowth,probs = c(.025)), 2) `, `r  round(quantile(correlation_bic$cor_sgrowth,probs = c(.975)), 2) `] or at Metis (median correlation with N [95% CRI]: `r  round(quantile(correlation_Metis$cor_sN,probs = c(.5)), 2) ` [`r  round(quantile(correlation_Metis$cor_sN,probs = c(.025)), 2) `, `r  round(quantile(correlation_Metis$cor_sN,probs = c(.975)), 2) `], median correlation with growth rate [95% CRI]: `r  round(quantile(correlation_Metis$cor_sgrowth,probs = c(.5)), 2) ` [`r  round(quantile(correlation_Metis$cor_sgrowth,probs = c(.025)), 2) `, `r  round(quantile(correlation_Metis$cor_sgrowth,probs = c(.975)), 2) `]. 



# Tables and figures{-}

```{r t1, echo=FALSE, verbose = FALSE}
#sitesnames<-data.frame(year = c(1998:2003, 2008:2016, 2019), file_name=NA, WAIC = NA)
s.sizes <- cbind(s.sizes.b, s.sizes.m[, -1])

s.sizes <- s.sizes[, c("year","min","max","unique.ID", "recap.b[, 4]", "nb.male", "nb.female","min","max","unique.ID" , "recap.m[, 4]", "nb.male", "nb.female")]
s.sizes <- s.sizes %>% mutate(across(where(is.numeric), round, 2))
knitr::kable(s.sizes,caption = "Sample sizes for the estimation of survival and abundance of harbour seal pups at Bic and Métis study sites, Québec, Canada. Number of pups, females and males are reported for each year and each site, with individuals of unknown sex were excluded (total of 8 individuals of unknown sex over the study period). ‘p.recapt’ is the proportion of individuals with >1 capture on the total number of marked individuals. No capture was conducted in years 2004-2007 and 2017-2018", "simple", col.names = c("Year", "Earliest capture", "Latest capture", "n.ind", "p.recapt", "n.males", "n.females", "Earliest capture", "Latest capture", "n.ind", "p.recapt", "n.males", "n.females"), align = "lcccccccccc")
```


## Figure{-}
Here figures are for model 3 in both cases but will have to check convergence issues first. 
```{r}
mytheme <- theme(plot.margin = unit(c(0.5,0.5,0.5,0.5), "cm"), legend.position = "none",
                 axis.text=element_text(size=12, color="white"),
                 axis.title=element_text(size=14, color="white"),
                 axis.ticks.x =element_line(color="white"),
                 axis.ticks.y = element_line(color="white"),
                 rect = element_rect(fill = "white"),
                 axis.line = element_line(color="white"),
                 panel.grid.major = element_line(color="white"),
                 panel.grid.minor = element_line(color="transparent"),
                 panel.background = element_rect(fill = "transparent",colour = NA),
                 plot.background = element_rect(fill = "transparent",colour = NA))
background_grid(
    major = c("xy", "x", "y", "only_minor", "none"),
    minor = c("none", "xy", "x", "y"),
    size.major = 0.5,
    size.minor = 0.2,
    color.major = "grey85",
    color.minor = "grey85"
)

```

```{r f1, echo=FALSE, fig.cap="Harbour seal pup abundance in the St. Lawrence estuary, 1998-2019, at Bic Island and Métis, Québec, Canada. Estimates are from a model with a fixed effect of sampling occasions on intra-annual survival and detection probability. Bars represent 95% credible intervals around posterior mean. Data were not available for years 2004-2007 and 2017-2018.", fig.width=7.165354, fig.height=5}
# results_all %>% 
#   ggplot(aes(x=yr, y=N, ymin=ymin, ymax=ymax, color=site)) +
#   geom_pointrange(position=position_dodge(w=0.5), size=0.75) +
#   labs(x=expression('Year')) + 
#   labs(y="Pup abundance") +  
#  # scale_y_continuous(limits = c(0,1),breaks = seq(from = 0, to = 1, by = 0.1)) +
#   scale_x_continuous(breaks = seq(from = 1998, to = 2019, by = 2)) +
#   #mytheme + background_grid()
#   theme(panel.grid.minor = element_line(color="transparent"))+
#   scale_colour_manual(values = c("bic" = "#FFDB6D", "metis" = "#00AFBB")) + 
#   theme_cowplot(12) +
#   theme(panel.grid.major = element_line(color="lightgrey", linetype = 'dotted'))


N_bic.3_summa$site <- 'bic'
N_metis.3_summa$site <- 'metis'
df_N <- rbind(N_bic.3_summa, N_metis.3_summa)

ggplot(df_N,aes(x=as.factor(yr), color = site))+
    geom_linerange(aes(ymin=Nsuper_cil,ymax=Nsuper_cih), position=position_dodge(w=0.75), size=0.75)+
    geom_linerange(aes(ymin=Nsuper_c25,ymax=Nsuper_c75),size=2, position=position_dodge(w=0.75), size=0.75)+
    geom_point(aes(y=Nsuper_mean),shape=18,size=4, position=position_dodge(w=0.75), size=0.75)+
    labs(x='Year',y='Pup abundance')+
  scale_y_continuous(limits = c(0,400),breaks = seq(from = 0, to = 400, by = 50)) +
  theme(panel.grid.minor = element_line(color="transparent"))+
  scale_colour_manual(values = c("bic" = "#FFDB6D", "metis" = "#00AFBB")) + theme_cowplot(10) +
  theme(panel.grid.major = element_line(color="lightgrey", linetype = 'dotted')) 
```


```{r f2, echo=FALSE, fig.cap="Pre-weaning survival of harbour seal in the St. Lawrence estuary, 1998-2019, at Bic Island and Métis, Québec, Canada. Estimates are from a model with a fixed effect of sampling occasions on intra-annual survival and detection probability. Bars represent 95% credible intervals around posterior mean. Data were not available for years 2004-2007 and 2017-2018. Two years did not converge: 2003 for Bic Island, and 2013 for Métis; the estimates are thus excluded from the figure.", fig.height=5, fig.width=7.165354}
# 
# results_weanSurv$yr <- as.numeric(as.character(results_weanSurv$yr))
# 
# results_weanSurv %>% 
#   ggplot(aes(x=yr, y=Mean, ymin=CIL, ymax=CIH, color=site)) +
#   geom_pointrange(position=position_dodge(w=0.75), size=0.75) +
#   labs(x=expression('Year')) + 
#   labs(y="Pre-weaning survival") +  
#   scale_y_continuous(limits = c(0,1),breaks = seq(from = 0, to = 1, by = 0.1)) +
#   scale_x_continuous(breaks = seq(from = 1998, to = 2019, by = 2)) +
#   theme(panel.grid.minor = element_line(color="transparent"))+
#   scale_colour_manual(values = c("bic" = "#FFDB6D", "metis" = "#00AFBB")) + theme_cowplot(12) +
#   theme(panel.grid.major = element_line(color="lightgrey", linetype = 'dotted')) 

# get means (june version)
meanSurvBic <- weanSurvOut_b_summa %>% filter(yr!=2003|site!='bic') 
meanSurvBic <- mean(meanSurvBic$weanSurv_mean)

meanSurvMetis <- weanSurvOut_m_summa %>% filter(yr!=2013|site!='metis')
meanSurvMetis <- mean(weanSurvOut_m_summa$weanSurv_mean)

# try wit this df 
weanSurvOut_b_summa$site <- 'bic'
weanSurvOut_m_summa$site <- 'metis'
df_weanSurv <- rbind(weanSurvOut_b_summa, weanSurvOut_m_summa)
df_weanSurv <- df_weanSurv %>% filter(yr!=2003|site!='bic') %>% filter(yr!=2013|site!='metis')

ggplot(df_weanSurv,aes(x=as.factor(yr), color = site))+
    geom_linerange(aes(ymin=weanSurv_cil,ymax=weanSurv_cih), position=position_dodge(w=0.75), size=0.75)+
    geom_linerange(aes(ymin=weanSurv_c25,ymax=weanSurv_c75),size=2, position=position_dodge(w=0.75), size=0.75)+
    geom_point(aes(y=weanSurv_mean),shape=18,size=4, position=position_dodge(w=0.75), size=0.75)+
    labs(x='Year',y='Pre-weaning survival')+
    scale_y_continuous(limits = c(0,1),breaks = seq(from = 0, to = 1, by = 0.1)) +
  theme(panel.grid.minor = element_line(color="transparent"))+
  scale_colour_manual(values = c("bic" = "#FFDB6D", "metis" = "#00AFBB")) + theme_cowplot(10) +
  theme(panel.grid.major = element_line(color="lightgrey", linetype = 'dotted')) 

  
```



```{r f3, echo=FALSE, fig.cap="Capture probability over time in harbour seal pups, Bic Island, Quebec, Canada, in a year with a) low sample size and recapture rate (2019, N = 56, recapture rate 0.05 and b) high sample size and recapture rate (year 1999, N=37, recapture rate = 0.81). full description of sample size per year appears in Table 1. ", fig.height=5, fig.width=7.165354}
cowplot::plot_grid(p_panel[[16]],p_panel[[2]], labels=c('a)','b)'))
```



# Supplementary Information

```{r JS model description, echo=FALSE}
jsmodels <- aicT[1:5, 1:2]
jsmodels$details <-NA

names(jsmodels) <- c('model', 'notation', 'description')
jsmodels[1,3]<-"A constant detection probability p and survival probability φ, across all individuals and occasions."
jsmodels[2,3]<-"A fixed time effect to account for within-season variation in p and a constant φ over the season. As pups grow, they are less easily captured (Dubé et al. 2003)."
jsmodels[3,3]<-"A fixed time effect on detection probability p, and a fixed time effect on survival probability φ."
jsmodels[4,3]<-"A time effect on φ but a constant capture probability, p."
jsmodels[5,3]<-"A random effect of time on both p and φ if individuals do not have the same capture or survival probability across the season because of local environmental conditions affecting swimming behavior, for example."

knitr::kable(jsmodels,caption = "Description of the Jolly-Seber models for the estimation of harbour seal pup abundance in the St. Lawrence Estuary, Quebec, Canada, 1998-2019. p: capture probability, b: recruitment, φ: survival probability, (.): a constant effect over sampling occasions, (t): a fixed time effects over sampling occasions, (1|t) : random time effects (in survival or capture probabilities). All models assumed fixed time effects in entry probabilities, b.", "simple")
```



```{r CJS model description, echo=FALSE}
cjsmodels <- waic.table[1:6, 1:2]
cjsmodels$details <-NA

names(cjsmodels) <- c('model', 'notation', 'description')


cjsmodels[1,3]<-"Constant detection probability p and pre-weaning survival probability φ, across all pups and occasions."
cjsmodels[2,3]<-"A fixed time effect to account for within-season variation in p, and a constant φ over the season."
cjsmodels[3,3]<-"A fixed time effect on detection probability p, and a random time effect on survival probability φ."
cjsmodels[4,3]<-"A constant capture probability p, and a random time effect on survival probability φ if pups do not have the same survival probability across the season."
cjsmodels[5,3]<-"A random individual effect on p to model capture heterogeneity, and a constant φ across the season."
cjsmodels[6,3]<-"A random individual effect on p to model capture heterogeneity, and a random time effect on φ if pups do not have the same survival probability across the season."

knitr::kable(cjsmodels,caption = "Description of the Cormack-Jolly-Seber models for the estimation of harbour seal pre-weaning survival in the St. Lawrence Estuary, Quebec, Canada, 1998-2019. p: capture probability, φ: pre-weaning survival probability at 30 days, (.): a constant effect over sampling occasions, (t): a fixed time effects over sampling occasions, (1|t) : random time effects (in survival or capture probabilities). All models assumed fixed time effects in entry probabilities, b.", "simple")
```

```{r JS model selection for both sites, echo=FALSE}
names <- c(
  ' ', "$p(.)\\phi(.)b(t)$","$p(t)\\phi(.)b(t)$", "$p(t)\\phi(t)b(t)$", "$p(.)\\phi(t)b(t)$", "$p(1|t)\\phi(1|t)b(t)$"
)
long_N_dt_b <- rbind(names, long_N_dt_b)
long_N_dt_m <- rbind(names, long_N_dt_m)

names2 <- c('','Bic Island',  rep('', 4), 'Metis', rep('', 4))
modSel_JS <- cbind(long_N_dt_b, long_N_dt_m[, 2:6])
modSel_JS <- rbind(names2, modSel_JS)

knitr::kable(modSel_JS,caption = "Jolly-Seber model selection for the estimation of pup abundance in a harbour seal population studied at Bic Island and Métis, Quebec, Canada, 1998-2019. Each year was modeled independently thus yielding a WAIC per year; the best model was selected based on number of years it yielded the lowest WAIC. p: capture probability, b: recruitment, φ: survival probability, (.): a constant effect over sampling occasions, (t): a fixed time effects over sampling occasions, (1|t) : temporal random effects (in survival and capture probabilities).", "simple")
```


```{r ts3}
N_bic.3_summa <- N_bic.3_summa %>% mutate(across(where(is.numeric), round, 2))
knitr::kable(N_bic.3_summa,caption = "Harbour seal pup abundance in the St. Lawrence Estuary, 1998-2019, at Bic Island, Québec, Canada. Estimates are from a model with a fixed effect of sampling occasions on intra-annual survival and detection probability. Mean posterior estimates, 25th and 75th quantiles, and 95% credible intervals around posterior means were extracted. Data were not available for years 2004-2007 and 2017-2018.", "simple")
```

```{r}
N_metis.3_summa <- N_metis.3_summa %>% mutate(across(where(is.numeric), round, 2))
knitr::kable(N_metis.3_summa,caption = "Harbour seal pup abundance in the St. Lawrence Estuary, 1998-2019, at Métis, Québec, Canada. Estimates are from a model with a fixed effect of sampling occasions on intra-annual survival and detection probability. Mean posterior estimates, 25th and 75th quantiles, and 95% credible intervals around posterior means were extracted. Data were not available for years 2004-2007 and 2017-2018.", "simple")
```


```{r ts4, echo=FALSE, verbose=FALSE}
names <- c(' ', "$p(.)\\phi(.)$", "$p(t)\\phi(.)$", "$p(t)\\phi(1|t)$", "$p(.)\\phi(1|t)$", "$p(1|i)\\phi(.)$", "$p(1|i)\\phi(1|t)$" )
long_Surv_dt_b <- rbind(names, long_Surv_dt_b)
knitr::kable(long_Surv_dt_b,caption = "Cormack-Jolly-Seber model selection for the estimation of pre-weaning survival in a harbour seal pups, Bic Island, Quebec, Canada. Each year was modeled independently thus yielding a WAIC per year; the best model was selected based on number of years it yielded the lowest WAIC. p: capture probability, phi : survival probability, (.) : a constant effect over sampling occasions, (t) : a fixed time effects over sampling occasions, (1|t) : temporal random effects (in survival and capture probabilities), (1|i) : individual random effects (capture probabilities).", "simple")
```


```{r ts5, echo=FALSE, verbose=FALSE}
long_Surv_dt_m <- rbind(names, long_Surv_dt_m)
knitr::kable(long_Surv_dt_m,caption = "Cormack-Jolly-Seber model selection for the estimation of pre-weaning survival in a harbour seal pups, Métis, Quebec, Canada. Each year was modeled independently thus yielding a WAIC per year; the best model was selected based on number of years it yielded the lowest WAIC. p: capture probability, phi : survival probability, (.) : a constant effect over sampling occasions, (t) : a fixed time effects over sampling occasions, (1|t) : temporal random effects (in survival and capture probabilities), (1|i) : individual random effects (capture probabilities).", "simple")
```


```{r}
weanSurvOut_b_summa <- weanSurvOut_b_summa %>% mutate(across(where(is.numeric), round, 2))
knitr::kable(weanSurvOut_b_summa,caption = "Harbour seal pup pre-weaning survival in the St. Lawrence Estuary, 1998-2019, at Bic Island, Québec, Canada. Estimates are from a model with constant (mean) intra-annual survival and fixed time effects on detection probability. Mean posterior estimates, 25th and 75th quantiles, and 95% credible intervals around posterior means were extracted. Data were not available for years 2004-2007 and 2017-2018.", "simple")
```

```{r}
weanSurvOut_m_summa <- weanSurvOut_m_summa %>% mutate(across(where(is.numeric), round, 2))
knitr::kable(weanSurvOut_m_summa,caption = "Harbour seal pup abundance in the St. Lawrence Estuary, 1998-2019, at Métis, Québec, Canada. Estimates are from a model with random time effects on intra-annual survival and fixed time effects on detection probability. Mean posterior estimates, 25th and 75th quantiles, and 95% credible intervals around posterior means were extracted. Data were not available for years 2004-2007 and 2017-2018.", "simple")
```

```{r echo=FALSE, verbose=FALSE}
# reorganise a bit 
colnames(meanBd_bic)
meanBd_bic <- meanBd_bic[, c("yr" , "Mean","Median","St.Dev.", "CIL","CIH")]
names(meanBd_bic) <- c("Year" , "MeanDate","MedianDate","St.Dev.", "Lower CRI","Upper CRI")

knitr::kable(meanBd_bic,caption = "Estimated parturition date of harbour seal pups, Bic Islanc, Quebec, Canada, 1998-2019. Parturition date was modeled as a function of body mass within a Cormack-Jolly-Seber model of pre-weaning survival. Each year was modeled independently and estimates are from a model with a fixed time effect on capture probability and a constant daily survival probability over time. No data were available for 2004-2007 and 2017-2018. CRI: 95% credibilibity interval.", "simple")
```
```{r echo=FALSE, verbose=FALSE}
# reorganise a bit 
colnames(meanBd_metis)
meanBd_metis <- meanBd_metis[, c("yr" , "Mean","Median","St.Dev.", "CIL","CIH")]
names(meanBd_metis) <- c("Year" , "Mean date","Median date","St.Dev.", "Lower CRI","Upper CRI")

knitr::kable(meanBd_metis,caption = "Estimated parturition date of harbour seal pups, Metis, Quebec, Canada, 1998-2019. Parturition date was modeled as a function of body mass within a Cormack-Jolly-Seber model of pre-weaning survival. Each year was modeled independently and estimates are from a model with a fixed time effect on capture probability and a random time effect on daily survival probability. No data were available for 2004-2007 and 2017-2018. CRI: 95% credibilibity interval.", "simple")
```





